When are Fuzzy Extractors Possible?

### 2015-02-09 info-bounds/FuzzyExtPossibility.tex
working on proof of leveled hashing

@@

\section{Proof of \thref{thm:layered hashing}}
\label{sec:proof of layered hashing}
\begin{proof}
Throughout the proof we assume that $\ell = n$ is the number of levels.  The proof can be carried out for an arbitrary $\ell$ but it leads to a complicated theorem statement.
We extend our notation for the maximum likelihood ball to the leveled case.  Define $\beta_{t,i}$ as the maximum number of points in a ball in level $i$.  That is,
\[
\beta_{t,i} = \max_{w' \in \mathcal{M}} \left|\{w | \dis(w, w')\le t \wedge \Pr[W=w]\in L_i\}\right|.
\]

\noindent
\textbf{Correctness:}  Fix some $w, w'$.  If $\Pr[W=w]\le 2^{-(m+\ell)} = 2^{-(m+n)}$, then $w$ is simply transmitted to $\rec$ and correctness is clear.  When $\Pr[W=w]> 2^{-(m+n)}$ let $L_i^*$ be the level of $\Pr[W=w]$.

@@

@@

In the standard secure sketch game, to maximize probability of predicting $w$, the adversary supplies the $w'$ with the most nearby probability mass.  With the additional information about the probability of $w$, the adversary should supply $w'$ with the most nearby probability in the correct probability level.  That is, the adversary's best guess is:
\[
w^*= \argmax_{w' \in \mathcal{M}}\sum_{w\in W | \Pr[W=w]\in L_i\wedge \dis(w, w')}\Pr[W=w].
\] 
The success of this adversary is at most $2^{-i}\beta_{t,i}$ as there at $\beta_{t,i}$ nearby points in that layer each with probability at most $2^{-i}$.  There are $n$ outcomes for $i$. The overall success of such an adversary is at most $n$ better than an adversary without such input~(by~\cite[Lemma 2.2]{DBLP:journals/siamcomp/DodisORS08}).  That is, 
\begin{align}
\expe_{i | m\le i \le m+n}2^{-i}\beta_{t,i}%&\le \expe_{i | m\le i \le n+m}\left( \max_{w^*\in W}\sum_{w\in W|  w\in L_i \wedge \dis(w, w^*) \le t}\Pr [W=w]\right) \nonumber\\
%&\le n \left(\max_{w^*\in W} \sum_{w\in W | \dis(w, w^*)\le t} \Pr[W=w]\right)\nonumber \\
&\ge n 2^{-\Hfuzz(W)}\label{eq:link fuzz 3}
\end{align}
\textbf{Security:}
%We just saw that an adversary that learns $L_i$ performs at most a factor of $n$ better than an adversary without this information.  
We now argue security of the construction.  First note that the total weight of points whose probability is less than $2^{-(n+m)}$ is at most $2^{-m}$~(there are at most $2^n$ points in the distribution).  Let $1_{\text{low}}$ be the indicator random variable for $\Pr[W=w]\le 2^{-(n+m)}$.  Then 
\begin{align}
\Hav(W | \sketch(W)) &= -\log \left(\Pr[1_{\text{low}}=1] * 1 + \Pr[1_{\text{low}} =0]   2^{-\Hav(W | \sketch(W) \wedge 1_{\text{low}} = 0)}\right)\nonumber\\
&\ge -\log\left( 2^{-m} + (1-2^{-m})2^{-\Hav(W | \sketch(W) \wedge 1_{\text{low}} = 0)}\label{eq:min-entropy}\right)
\end{align}
For the remainder of the proof, we seek a bound on $\Hav(W | \sketch(W) \wedge 1_{\text{low}} =0).$

@@

\begin{lemma}
\label{lem:fuzz can't get key}
Fix some $(\gen, \rep)$ algorithm with $\kappa \ge 2$.  There exists an information-theoretic distinguisher between $(R, P, Z)$ and $(U_\kappa, P, Z)$ with advantage $\epsilon = 1/8-\ngl(n)$.
\end{lemma}
\begin{proof}
As in the proof of \thref{thm:imposs sketch}, we assume that $\rep$ is deterministic.  Denote by $(\Key, P) \leftarrow \gen(V)$.
By Markov's inequality, there exists a set $A_{p}$ such that $\Pr[p\in A_{p}]\ge 1/2$ and $\forall p\in A_{p}$, 
\[
(\Key |P =p, P = p ) \approx_{2\epsilon} (U_\kappa , P =p).
\]
%\{w' | \dis (w', w)\le t \wedge \rec(w', p) \neq w\}\le 2\delta < 1/2.\]

@@

\begin{lemma}
$|\inter(Q_{p^*, \key^*})| \le 2^{n-4\nu}$.
\end{lemma}
\begin{proof}
By the isoperimetric inequality on the Hamming space~(we use a version due to~\cite[Theorem 1]{frankl1981short}, the original result is due to Harper~\cite{harper1966optimal}), there exists a $\eta$-deficient ball $S_{p^*, \key^*}$ centered at $0$ and a set $D$ such that $|S_{p^*, \key^*}| = |\inter(Q_{p^*, \key^*})|$, $|D| = |Q_{p^*, \key^*}^\complement|$ and $\forall s\in S_{p^*, \key^*}, d\in D$, $\dis(s, d) \ge t$~(alternatively, the distance between the sets is $t$).  Furthermore, note that $S_{p^*, \key^*} \cup D$ is a deficient ball~(and its radius is $\eta+t$).
We now find bound the size of $S_{p^*, \key^*}$.

Recall that $|S_{p^*, \key^*} \cup D| = |Q_{p^*, \key^*} | \le 2^{n-\kappa}\leq |\mathcal{M}|/2$.  Since this set contains less than half the points in the metric space we know its radius at most $n/2$.  This means that $|S_{p^*, \key^*}|$ is a deficient sphere of radius at most $n/2-t$.  Let $X$ denote a uniform string on $\zo^n$.  We use Hoeffding's inequality~\cite{hoeffding1963probability}:

\begin{align*}
|S_{p^*, \key^*}| \le \{ x | \dis (x, 0)\le n-t\} &= 2^n \Pr_{X\leftarrow \zo^n} [ wt(X) \le (1/2-t/n)n] \le 2^n e^{-n ((t/n)^2)} = 2^n e^{-4\nu} \le 2^{n - 4\nu}
\end{align*}
\end{proof}

>>>>

Before proving security of \consref{cons:leveling} we extend the notation to the maximum likelihood ball containing points of a given probability.  Define $\beta_{t,i}$ as the maximum number of points in a ball in level $i$.  That is,
\[
\beta_{t,i} = \max_{w' \in \mathcal{M}} \left|\{w | \dis(w, w')\le t \wedge \Pr[W=w]\in L_i\}\right|.
\]

@@

\section{Proof of \thref{thm:layered hashing}}
\label{sec:proof of layered hashing}
\begin{proof}
Throughout the proof we assume that $\ell = n$ is the number of levels.  The proof can be carried out for an arbitrary $\ell$ but it leads to a complicated theorem statement.

\paragraph{Correctness:}  Fix some $w, w'$.  If $\Pr[W=w]\le 2^{-(m+\ell)} = 2^{-(m+n)}$, then $w$ is simply transmitted to $\rec$ and correctness is clear.  When $\Pr[W=w]> 2^{-(m+n)}$ let $L_i^*$ be the level of $\Pr[W=w]$.

@@

\paragraph{Security:}
We now argue security of the construction.  First note that the total weight of points whose probability is less than $2^{-(n+m)}$ is at most $2^{-m}$~(there are at most $2^n$ points in the distribution).  Let $1_{\text{low}}$ be the indicator random variable for $\Pr[W=w]\le 2^{-(n+m)}$.  Then 
\begin{align}
\Hav(W | \sketch(W)) &= -\log \left(\Pr[1_{\text{low}}=1] * 1 + \Pr[1_{\text{low}} =0]   2^{-\Hav(W | \sketch(W) \wedge 1_{\text{low}} = 0)}\right)\nonumber\\
&\ge -\log\left( 2^{-m} + (1-2^{-m})2^{-\Hav(W | \sketch(W) \wedge 1_{\text{low}} = 0)}\label{eq:min-entropy}\right)
\end{align}
For the remainder of the proof, we seek a bound on $\Hav(W | \sketch(W) \wedge 1_{\text{low}} =0).$

We now show that the fuzzy min-entropy conditioned on the level information is proportional to $\beta_{t, i}$.  Let $I$ be the random variable of the level information.
\begin{claim}
$\Hoo(W | I=i ) -\log \beta_{t, i} \ge \Hfuzz(W | I =i ) \ge \Hoo(W | I =i ) -\log \beta_{t, i}-1$.
\end{claim}
\begin{proof}
We show the left inequality first.
\begin{align*}
\Hfuzz(W|I=i) &= -\log \left(\max_{w' \in \mathcal{M}} \left( \sum_{w\in W | \Pr[W=w]\in L_i \wedge \dis(w, w')} \Pr[W=w | I=i]\right) \right)\\
&\le -\log \left(\max_{w' \in \mathcal{M}} \left( \sum_{w\in W | \Pr[W=w]\in L_i \wedge \dis(w, w')} 2^{-(m+i)}\right) \right)\\
&= (m+i)- \log\left(\max_{w' \in \mathcal{M}} \left( \sum_{w\in W | \Pr[W=w]\in L_i \wedge \dis(w, w')} 2^{-(m+i)}\right) \right)\\
&=(m+i)- \log\left(\max_{w' \in \mathcal{M}} \left( \sum_{w\in W | \Pr[W=w]\in L_i \wedge \dis(w, w')} 2^{-(m+i)}\right) \right)\\
&=(m+i)-  \log \left(\max_{w' \in \mathcal{M}} \left|\{w | \dis(w, w')\le t \wedge \Pr[W=w]\in L_i\}\right|\right)\\
&= m+i-\log \beta_{t, i}.
\end{align*}
We now show the right inequality.
\end{proof}

@@

In the standard secure sketch game, to maximize probability of predicting $w$, the adversary supplies the $w'$ with the most nearby probability mass.  With the additional information about the probability of $w$, the adversary should supply $w'$ with the most nearby probability in the correct probability level.  That is, the adversary's best guess is:
\[
w^*= \argmax_{w' \in \mathcal{M}}\sum_{w\in W | \Pr[W=w]\in L_i\wedge \dis(w, w')}\Pr[W=w].
\] 
The success of this adversary is at most $2^{-i}\beta_{t,i}$ as there at $\beta_{t,i}$ nearby points in that layer each with probability at most $2^{-i}$.  There are $n$ outcomes for $i$. The overall success of such an adversary is at most $n$ better than an adversary without such input~(by~\cite[Lemma 2.2]{DBLP:journals/siamcomp/DodisORS08}).  That is, 
\begin{align}
\expe_{i | m\le i \le m+n}2^{-i}\beta_{t,i}%&\le \expe_{i | m\le i \le n+m}\left( \max_{w^*\in W}\sum_{w\in W|  w\in L_i \wedge \dis(w, w^*) \le t}\Pr [W=w]\right) \nonumber\\
%&\le n \left(\max_{w^*\in W} \sum_{w\in W | \dis(w, w^*)\le t} \Pr[W=w]\right)\nonumber \\
&\ge n 2^{-\Hfuzz(W)}\label{eq:link fuzz 3}
\end{align}
\textbf{Security:}
%We just saw that an adversary that learns $L_i$ performs at most a factor of $n$ better than an adversary without this information.

@@

\begin{lemma}
\label{lem:fuzz can't get key}
Fix some $(\gen, \rep)$ algorithm with $\kappa \ge 2$.  There exists an information-theoretic distinguisher between $(R, P, Z)$ and $(U_\kappa, P, Z)$ with advantage $\epsilon = 1/8-\ngl(n)$.
\end{lemma}
\begin{proof}
As in the proof of \thref{thm:imposs sketch}, we assume that $\rep$ is deterministic.  Denote by $(\Key, P) \leftarrow \gen(V)$.
By Markov's inequality, there exists a set $A_{\epsilon}$ such that $\Pr[p\in A_{\epsilon}]\ge 1/2$ and $\forall p\in A_{\epsilon}$, 
\[
(\Key |P =p) \approx_{2\epsilon} (U_\kappa , P =p).
\]
%\{w' | \dis (w', w)\le t \wedge \rec(w', p) \neq w\}\le 2\delta < 1/2.\]

@@

\begin{lemma}
$|\inter(Q_{p^*, \key^*})| \le 2^{n-4\nu}$.
\end{lemma}
\begin{proof}
By the isoperimetric inequality on the Hamming space~(we use a version
due to~\cite[Theorem 1]{frankl1981short}, the original result is due
to Harper~\cite{harper1966optimal}), there exists a $\eta$-deficient
ball $S_{p^*, \key^*}$ centered at $0^n$ and a deficient ball $D$,
centered at $1^n$, such that
$|S_{p^*, \key^*}| = |\inter(Q_{p^*, \key^*})|$, $|D| = |Q_{p^*,
  \key^*}^\complement|$ (where $\cdot^\complement$ denotes the
complement of a set) and $\forall s\in S_{p^*, \key^*}, d\in D$,
$\dis(s, d) \ge t$~(that is, the distance between the sets is $t$).  Note that $S_{p^*, \key^*} \cup D^\complement$ is a deficient ball~(and its radius is $\eta+t$).\lnote{Why?}
We now bound the size of $S_{p^*, \key^*}$.

Recall that $|S_{p^*, \key^*} \cup D| = |Q_{p^*, \key^*} | \le 2^{n-\kappa}\leq |\mathcal{M}|/2$.  Since this set contains less than half the points in the metric space we know its radius at most $n/2$.  This means that $|S_{p^*, \key^*}|$ is a deficient sphere \lnote{ball} of radius at most $n/2-t$.  Let $X$ denote a uniform string on $\zo^n$.  We use Hoeffding's inequality~\cite{hoeffding1963probability}: \lnote{should the first $n$ in the line below be $n/2$?}

\begin{align*}
|S_{p^*, \key^*}| \le \{ x | \dis (x, 0)\le n-t\} &= 2^n \Pr_{X\leftarrow \zo^n} [ wt(X) \le (1/2-t/n)n] \le 2^n e^{-n ((t/n)^2)} = 2^n e^{-4\nu} \le 2^{n - 4\nu}
\end{align*}
where the second to last equality follows from the definition of $t$
(as $4\nu \sqrt{n}$) at the beginning of the proof of Theorem~\ref{thm:imposs fuzz ext}.
\end{proof}
